---
title: 'Time Series Stylometry on the #TravelBan News Writers'
author: "Kanishka Misra"
date: "April 20, 2018"
output: 
  md_document: default
  pdf_document: default
---

## Setup

loading libraries

```{r warning = F, message = F}
library(tidyverse)
library(jsonlite)
library(tidyjson)
library(widyr)
library(tidytext)
library(ggridges)
library(lubridate)

options(scipen = 99)
```


## Loading data

```{r}
news_articles <- fromJSON("data/1996_dicts_9_sources.json", simplifyDataFrame = T) %>%
  as.tibble()

news_articles
```

## Tidying and Nesting

1. Count authors, only keep single author news articles.
2. Remove duplicates, and keep only text based duplicates if source is differenct.
3. Remove columns that are not relevant to the scope of this project.
4. Nest based on author, store article count.
5. Select top 10 (or more if similar number of articles exist) authors based on article count.

```{r}
nested_articles <- news_articles %>%
  distinct(text, source, .keep_all = T) %>%
  mutate(length = map_int(authors, length)) %>%
  filter(length < 2) %>%
  unnest(authors) %>%
  select(text, author = authors, publish_date, source) %>%
  group_by(author) %>%
  nest() %>%
  mutate(article_count = map_int(data, nrow)) %>%
  top_n(10, article_count)
  
nested_articles
```

We now have a tibble of the 12 authors with the most number of articles. We can then write a generic function that maps over each author and then calculates the distances of each authors news articles from their first article during the travel ban proposal.

```{r}
temporal_delta <- function(df) {
  df %>%
    mutate(
      text = str_replace_all(tolower(text), paste0(source, "|story highlights"), "")
    ) %>%
    select(-source) %>%
    group_by(publish_date) %>%
    unnest_tokens(word, text) %>%
    count(word) %>%
    ungroup() %>%
    mutate(
      publish_date = lubridate::ymd(publish_date)
    ) %>%
    pairwise_delta(publish_date, word, n, upper = F) %>%
    filter(item1 == min(item1)) %>%
    transmute(
      publish_date = item2,
      delta
    )
}

temporal_delta(nested_articles[1, ]$data[[1]])
```

If we map this to all writer's data, we get the following plot:

```{r}
author_deltas <- nested_articles %>%
  mutate(ts_delta = map(data, temporal_delta)) %>%
  unnest(ts_delta)

author_deltas %>%
  ggplot(aes(publish_date, delta)) + 
  geom_line() + 
  facet_wrap(~author)
```

Help me.
